Analysis

### 2017-03-26 Analysis.Rmd
Run several iterations of the models for better random errors
Add new custom functions for graphing

Add new diagnostic graphs for random errors

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

## 2015 Leaf Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2015 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2015),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

@@

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_lmm_2015$Sol[, c(1, 7:10)])
reps <- melt(reps)

@@

# Create a dummy x-axis variable for plotting
x <- 1:nrow(LShB_lmm_2015$Sol)

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatements
ggplot(data = trts, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(LShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)
```

## 2015 Tiller Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_lmm_2015 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2015),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

@@

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatments
ggplot(data = trts, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(TShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)
```

@@

## 2016 Leaf Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2016 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2016),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

@@

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatments
ggplot(data = trts, aes(x = rep(x, 3), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(TShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)

```

## Tiller Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_lmm_2016 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE, ~REP, 
                 data = as.data.frame(AUDPS_2016),
                 verbose = FALSE,
                 prior = eprior,
                 pr = TRUE)

@@

# Diagnostic line plots for replicate
ggplot(data = reps, aes(x = rep(x, 5), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for replicate
ggplot(reps[reps$variable != "X.Intercept.", ], aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# Diagnostic line plots for treatments
ggplot(data = trts, aes(x = rep(x, 3), y = value, group = variable)) +
  geom_line() +
  theme_tufte() +
  xlab("Iteration Number") +
  ylab(NULL) +
  facet_grid(variable ~ .)

# Posterior distributions for treatment
ggplot(trts, aes(x = value, color = variable)) + 
  geom_density() +
  theme_tufte()

# joint distibution of error
rdf <- data.frame(TShB_lmm_2015$VCV)
ggplot(rdf, aes(x = sqrt(REP), y = sqrt(units))) + 
  geom_density2d() + 
  geom_abline(intercept = 0, slope = 1)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```
## Notes on this analysis before starting

#### Changes between 2015 and 2016

Before the analysis, note that due to changes between the years, the analysis
must be carried out on each year separately. The 2015 data and 2016 data cannot
be combined due to changes in inoculation methods; there are other changes too,
but the main one is this. Therefore, the analyses will be conducted separately for
leaf severity and Leaf severity for 2015 dry season and 2016 dry season.
Comparisons will only be observational and cannot be statistically compared.

#### Random effects
Fitting the models with the default iterations resulted in some less than
acceptable models. Because of this, the number of iterations has been increased,
`nitt = 5e+05` and `burnin = 5000` and `thin = 100`. This results in much
smaller errors for random effects and better models.

## 2015
### 2015 Leaf Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2015 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE, ~REP, 
                          data = as.data.frame(AUDPS_2015),
                          verbose = FALSE,
                          prior = eprior,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          pr = TRUE)

@@

# create data frames for generating diagnostic plots
reps <- data.frame(LShB_lmm_2015$Sol[, c(1, 7:10)])
x <- 1:nrow(reps)
reps <- melt(reps)

@@

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2015 Diagnostic Plots for Leaf ShB Reps")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2015 Replicate Posteriors for Leaf ShB Reps")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2015 Diagnostic Plots for Leaf ShB Treatments")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2015 Treatment Posteriors for Leaf ShB Treatments")

# check random effects
plotTrace(LShB_lmm_2015$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(LShB_lmm_2015$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2015 Leaf ShB Random Error Distribution"))
```

### 2015 Tiller Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_lmm_2015 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE, ~REP, 
                          data = as.data.frame(AUDPS_2015),
                          verbose = FALSE,
                          prior = eprior,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          pr = TRUE)

@@

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2015 Diagnostic Plots for Leaf ShB Reps")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2015 Replicate Posteriors for Leaf ShB Reps")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2015 Diagnostic Plots for Leaf ShB Treatments")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2015 Treatment Posteriors for Leaf ShB Treatments")

# check random effects
plotTrace(TShB_lmm_2015$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_lmm_2015$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2015 Leaf ShB Random Error Distribution"))
```

@@

## 2016

### 2016 Leaf Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
LShB_lmm_2016 <- MCMCglmm(LShB_AUDPS ~ WMGT * NRTE, ~REP, 
                          data = as.data.frame(AUDPS_2016),
                          verbose = FALSE,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          prior = eprior,
                          pr = TRUE)

@@

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Leaf ShB Reps")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Replicate Posteriors for Leaf ShB Reps")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2016 Diagnostic Plots for Leaf ShB Treatments")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Treatment Posteriors for Leaf ShB Treatments")

# check random effects
plotTrace(LShB_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(LShB_lmm_2015$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2016 Leaf ShB Random Error Distribution"))
```

### 2016 Tiller Sheath Blight Severity Model
```{r}

eprior <- list(R = list(V = 1, nu = 0.02),
               G = list(G1 = list(V = 1, nu = 0.02, alpha.V = 1000)))
TShB_lmm_2016 <- MCMCglmm(TShB_AUDPS ~ WMGT * NRTE, ~REP, 
                          data = as.data.frame(AUDPS_2016),
                          verbose = FALSE,
                          nitt = 5e+05,
                          burnin = 5000,
                          thin = 100,
                          prior = eprior,
                          pr = TRUE)

@@

# diagnostic line plots for replicate
plot_diagnostic_lines(d = reps,
                      x = x,
                      title = "2016 Diagnostic Plots for Leaf ShB Reps")

# posterior distributions for replicate
plot_replicate_posteriors(d = reps,
                          title = "2016 Replicate Posteriors for Leaf ShB Reps")

# diagnostic line plots for treatments
plot_diagnostic_lines(d = trts,
                      x = x,
                      title = "2106 Diagnostic Plots for Leaf ShB Treatments")

# Posterior distributions for treatment
plot_treatment_posteriors(d = trts,
                          title = "2016 Treatment Posteriors for Leaf ShB Treatments")

# check random effects
plotTrace(TShB_lmm_2016$VCV, log = TRUE)

# plot joint distibution of error
rdf <- data.frame(TShB_lmm_2016$VCV)
plot_joint_random_error_dist(d = rdf,
                             title = ("2016 Leaf ShB Random Error Distribution"))
```

# Conclusions

The models all appear to be good fits after increasing the number of iterations.
None of the treatments in either year, 2015 or 2016, were significant. The
see `pMCMC` values and also posterior graphs. There is a large amount of overlap
in all of the posterior graphs.
